% The documentation of the usage of CTUstyle -- the template for
% typessetting thesis by plain\TeX at CTU in Prague
% ---------------------------------------------------------------------
% Petr Olsak  Jan. 2013

% You can copy this file to your own file and do some changes.
% Then you can run:  pdfcsplain your-file

\input ctustyle  
\def\thednum{(\the\chapnum.\the\dnum)}
 % The template is included here.
%\input pdfuni    % Uncomment this if you need accented PDFoutlines
%\input opmac-bib % Uncomment this for direct reading of .bib database files 

\worktype [B/EN] % Type: B = bachelor, M = master, D = Ph.D., O = other
                 % / the language: CZ = Czech, SK = Slovak, EN = English

\faculty    {F3}  % Type your faculty F1, F2, F3, etc.
            % use main language of your document here:
\department {Department of Radioelectronics}
\title      {Distributed signal processing in~radio communication networks}
%\subtitle   {}
            % \subtitle is optional
\author     {Jakub Kolář}
\date       {May 2017}
\supervisor {prof. Ing. Jan Sýkora, CSc.}  % One or more supervisors
\studyinfo  {Open Electronic Systems}  % Study programme etc.
\workname   {Bachelor's thesis} % Used only if \worktype [O/*] (Other)
            % optional more information about the document:
%\workinfo   {kolarj39@fel.cvut.cz}
            % Title / Subtitle in minor language:
\titleEN    {CTUstyle -- the user manual}
\subtitleEN {the plain\TeX{} template for theses at CTU}
            % If minor language is other than English
            % use \titleCZ, \subtitleCZ or \titleSK, \subtitleSK instead it.
\pagetwo    {}  % The text printed on the page 2 at the bottom.

\abstractEN { The major objective of this Bachelor’s thesis
is to provide an introduction to Linear
average consensus algorithm on a graph. It is a distributed iterative algorithm,
that works with values assigned to
vertices of the graph and the goal is to
asymptotically obtain  an average value of the initial values in all vertices, respecting the topology of the graph. In the text, we have used a
matrix representation of the graph. To
design the parameters of the algorithm we
used the Laplacian matrix. Using matrix
representation, the algorithm that solves
the original problem is transferred to
iterative matrix multiplication. In the next step,
we have provided examples of the run of the algorithm in the case of ideal and reliable communication and also an outline of the solution
in the case of zero-mean noisy updates. Included are Matlab scripts
with the algorithm implementation. More
difficult formal proofs and derivations
are not included, but we have provided references, where it is possible to find them in the literature. In the final part, we've implemented a few simple examples of the described algorithm, motivated by radio communication.
}

\abstractCZ { Cílem této bakalářské práce je seznámení se s lineárním konsenzuálním algoritmem nad grafy. Jde o distribuovaný iterativní algoritmus, který pracuje s hodnotami přiřazenými vrcholům grafu a jehož cílem je, aby asymptoticky byly tyto hodnoty ve všech vrcholech stejné a průměrem hodnot počátečních. Studovaný algoritmus toho docílí jen komunikací mezi dvojicemi vrcholů přes existující hrany grafu.  V práci využíváme maticovou reprezentaci grafu a pro nalezení vhodných parametrů algoritmu používáme Laplaceovu matici. Algoritmus je tak převeden na opakující se maticové násobení, jehož význam zachovává původní úlohu. Dále uvádíme  názorné ukázky toho, jak algoritmus pracuje v případě ideálních aktualizací a základní řešení v případě ovlivňování aktualizací šumem s nulovou střední hodnotou. Součástí práce je implementace algoritmu v Matlabu. V práci nejsou uvedeny složitější důkazy a odvození, ale odkazujeme se na příslušnou uvedenou literaturu. Na závěr uvádíme několik jednoduchých příkladů využívajících popsané iterativní schéma.
}          

\keywordsEN {       Graph, Laplacian, Linear average consensus algorithm, Noisy updates, Estimation

}
\keywordsCZ {%
   Graf, Laplaceova matice, Lineární konsenzus, Aktualizace s rušením, Odhad parametru
}
\thanks {           %Use main language here
I would like to thank the supervisor of this thesis, prof.~Ing.~Jan Sýkora,~CSc., for such an interesting assignment. Also, I'd like to write here thanks to my beloved friends and family, who all inspire me and motivate me as well.
}
\declaration { 
I declare that I completed the presented thesis independently and that all used sources are quoted in accordance with the Methodological Instructions that cover the ethical principles for writing an academic thesis.

In Prague, 26. 5. 2017

\hfil\break\hfil\break\hfil\break\hfil\break

Prohlašuji, že jsem předloženou práci vypracoval samostatně a že jsem uvedl veškeré použité informační zdroje v~souladu s Metodickým pokynem o dodržování etických principů při přípravě vysokoškolských závěrečných prací.

V Praze, 26. 5. 2017

 % !!! Attention, you have to change this item.
   \signature % makes dots
}
\specification {\picw=\hsize \cinspic appendix/assignment.pdf}


%%%%% <--   % The place for your own macros is here.

%\draft     % Uncomment this if the version of your document is working only.
%\linespacing=1.7  % uncomment this if you need more spaces between lines
                   % Warning: this works only when \draft is activated!
%\savetoner        % Turns off the lightBlue backround of tables and
                   % verbatims, only for \draft version.
%\blackwhite       % Use this if you need really Black+White thesis.
%\onesideprinting  % Use this if you really don't use duplex printing. 

\makefront  % Mandatory command. Makes title page, acknowledgment, contents etc.
\input opmac-bib
\input introduction  % Files where the source of the document is prepared.
\input graph_theory
\input consensus_problem
\input distributed_estimation_in_wirelless_network
\input prilohy



\bye
